---
title: "Outbreak response models of vaccine-preventable diseases in humans (1970-2019) - A systematic review (Supplementary Files)"
author: "James Azam"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    keep_md: yes
    reference_docx: word-style.docx
    toc: yes
---

```{r echo=FALSE, message=FALSE, include=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('ggthemes')
library('scales')
library('janitor')
library('forcats')
library('stringr')
library('countrycode')
library('bib2df')
library('conflicted')

#resolve package function conflicts
conflict_prefer('filter', 'dplyr')
conflict_prefer('select', 'dplyr')
conflict_prefer('kable', 'knitr')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Sets the working directory for subsequent chunks, but not this chunk
#knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r echo=FALSE, include=FALSE}
# Load the cleaned data 
long_data <- read_rds('../data/final_data/cleaned_data/review_data_long_cleaned.rds')

#The cleaned data without citations
compact_data_init <- read_rds('../data/final_data/cleaned_data/review_data_compact_cleaned.rds')

compact_data <- compact_data_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

# The data with citations 
compact_data_with_citation_keys_init <- read_rds(normalizePath('../data/final_data/cleaned_data/compact_data_with_citation_keys_cleaned.rds')) 

compact_data_with_citation_keys <- compact_data_with_citation_keys_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

#Rearrange the columns
compact_data_with_citation_keys <- compact_data_with_citation_keys %>% 
  relocate(c(title, bibtexkey), .before = publication_year) %>% 
  relocate(publication_year, .before = bibtexkey) %>% 
  relocate(c(x0_reviewer, category, author:language), 
           .after = collab_type
           )

# Remove the FMD papers (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd') 

compact_data_no_fmd <- compact_data %>% filter(disease != 'fmd') 

compact_data_with_citation_keys_no_fmd <- compact_data_with_citation_keys %>% filter(disease != 'fmd') 
```
```{r echo=FALSE}
#' Count by collaboration types and make a bar plot
collab_types_count <- compact_data_no_fmd %>%
    count(author_affiliation_type, sort = TRUE) %>% 
    mutate(author_affiliation_type = as_factor(author_affiliation_type))

knitr::kable(collab_types_count %>% arrange(desc(n)))
```

### Trend in collaboration over the years

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
    
collap_types_count_barplot <- collab_types_count %>%
    ggplot() + geom_bar(aes(x = reorder(author_affiliation_type, n),
                            y = n
                            ),
                        stat = 'identity',
                        color = 'black'
                        ) +
    scale_y_continuous(breaks = seq(0, 140, 20),
                       labels = seq(0, 140, 20)
                       ) +
    labs(title = 'Number of studies per collaboration type', 
         x = '', 
         y = 'Number of studies') + 
  coord_flip() +
  theme_minimal() +
  theme(legend.position = 'none')

plot(collap_types_count_barplot)

knitr::kable(collab_types_count, 
             col.names = c('Collaboration type', 'Number of publications'),
             caption = 'Number of publications per collaboration type'
             )

```
# Number of studies per disease and intervention.

```{r echo=FALSE}
#count the number of unique interventions modelled
interventions_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(intervention_modelled, .keep_all = TRUE) %>% 
  ungroup()

#interventions are disease specific
interventions_per_disease <- interventions_unique_entries %>% 
  mutate(intervention_modelled = str_trim(intervention_modelled)) %>% 
  filter(intervention_modelled != 'e.g.') %>% #remove false rows
  filter(intervention_modelled != 'various') %>% #remove false rows
  group_by(disease) %>% 
  count(intervention_modelled, name = 'num_of_studies') %>% 
  arrange(disease, desc(num_of_studies)) %>% 
  ungroup()

knitr::kable(interventions_per_disease)

```

# Number of studies per outcome

```{r echo=FALSE}
#count the number of unique outcomes measured
outcomes_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(outcome_measured, .keep_all = TRUE) %>% 
  ungroup()
  
#Count the number of studies 
outcomes_count <- outcomes_unique_entries %>% 
  group_by(outcome_measured) %>% 
  count(outcome_measured, name = 'num_of_studies', sort = T) %>% 
  ungroup() %>% 
  filter(outcome_measured != 's') 

knitr::kable(outcomes_count)

#Plot the result
outcome_count_plot <- ggplot(data = outcomes_count) + 
  geom_point(aes(x = reorder(outcome_measured, 
                             num_of_studies), 
               y = num_of_studies
               ), 
           stat = 'identity',
           color = 'darkslategray4'
           ) +
  coord_flip() +
  theme_minimal()

plot(outcome_count_plot)


```

# Number of studies per model structure type

```{r echo=FALSE}
#Keep only the unique entries 
model_structure_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(model_structure) %>% 
  ungroup()

#Count the number of studies
model_structure_unique_count <- model_structure_unique_entries %>%
    group_by(model_structure) %>% 
    count(model_structure, name = 'num_of_studies') 

#Print the table
knitr::kable(model_structure_unique_count)
```

# Model structure, parametrization, and validation types

```{r echo=FALSE}
#Cross tally the model structure with parametrization and validation categories
#FIXME: Check that studies are either not being under counted or double-counted. What do I want to count here? Is it the unique studies per method or combined method because most studies used a combination of parametrization methods especially
parametrization_and_validation_by_model_structure <- compact_data_no_fmd %>%
  mutate(parametrization = str_replace(parametrization, ' ', ' + '),
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(model_structure, parametrization, validation) %>% 
  pivot_wider(names_from = model_structure, values_from = n)
  

#Print the table of results
knitr::kable(parametrization_and_validation_by_model_structure,
             caption = 'Parametrization and validation types per model structure')

```

# Model parametrization & validation

```{r echo=FALSE}
#Keep only the unique entries
parametrization_and_validation_count <- compact_data_no_fmd %>% 
  mutate(parametrization = str_replace(parametrization, ' ', ' + '), 
         validation = str_replace(validation, ' ', ' + ')
         ) %>% 
  count(parametrization, 
        validation, 
        name = 'num_of_studies'
        ) %>% 
  group_by(parametrization) %>% 
  arrange(desc(num_of_studies), .by_group = T)   

#Print the table
knitr::kable(parametrization_and_validation_count)

```

# How is sensitivity analysis distributed by model structure?

```{r echo=FALSE}
#Keep only the unique entries
sensitivity_analysis_unique_entries <- long_data_no_fmd %>% 
  group_by(title) %>% 
  distinct(sensitivity_analysis, .keep_all = TRUE) %>% 
  ungroup()

#Count the number of studies
sensitivity_analysis_count <- sensitivity_analysis_unique_entries %>% 
  count(model_structure, sensitivity_analysis, name = 'num_of_studies') 

#Print the table
knitr::kable(sensitivity_analysis_count)

```

# Data use and availability

```{r echo=FALSE}
#Keep only the unique entries
code_and_data_availability_count <- compact_data_no_fmd %>% 
  count(data_used, data_available) %>% 
  pivot_wider(names_from = data_available, values_from = n) %>% 
  replace_na(list(not_applicable = 0, 
                  yes = 0, 
                  no = 0
                  )
             )

#Print the results
knitr::kable(code_and_data_availability_count, caption = 'Data use and availability')

```
# Code availability

```{r echo=FALSE}
#Keep only the unique entries
code_availability_count <- compact_data_no_fmd %>% 
  count(simulation_code_available) 

#Print the results
knitr::kable(code_availability_count, caption = 'Code availability')

```

# Number of studies per disease 
## By outbreak type

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_outbreak_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         outbreak_type = as_factor(str_replace(outbreak_type, ' ', '+'))
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, outbreak_type, sort = TRUE) %>%
  pivot_wider(names_from = outbreak_type, values_from = n) %>% 
  replace_na(list(hypothetical_outbreak = 0, real_outbreak = 0)) %>% 
  mutate(total = hypothetical_outbreak + real_outbreak) %>% 
  arrange(desc(total))

knitr::kable(disease_studied_by_outbreak_type)
```

## By objective type

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_objective_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  mutate(disease = as_factor(disease), 
         objectives = as_factor(objectives)
         ) %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, objectives) %>%
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(future = 0, past = 0, both = 0)) %>% 
  mutate(total = future + past + both) %>% 
  arrange(desc(total))

knitr::kable(disease_studied_by_objective_type)
```

## FMD by collaboration type
```{r echo=FALSE}
fmd_studies_by_collab_type <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  mutate(disease = as_factor(disease), 
         author_affiliation_type = as_factor(str_replace(author_affiliation_type, ' ', '+'))
         ) %>% 
  mutate(collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         ) %>% 
  count(collab_type, sort = TRUE) 

knitr::kable(fmd_studies_by_collab_type)
```
## FMD by objective type
```{r echo=FALSE}
fmd_studies_by_objective_and_outbreak_type <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  count(objectives, outbreak_type) %>% 
  pivot_wider(names_from = outbreak_type, values_from = n)

knitr::kable(fmd_studies_by_objective_and_outbreak_type)
```


# Secondary summaries
## Studies before 2006 (low frequency period) 
```{r echo=FALSE}
#Number of studies before 2005 (low frequency period)
n_studies_before_2006 <- compact_data_no_fmd %>% 
  filter(publication_year < 2006) %>% 
  summarise(n = n())

print(paste('Number of studies before 2006:', n_studies_before_2006))
```
## Number of studies between 2006-2019 (high frequency period)
```{r echo=FALSE}
n_studies_btwn_2006_2019 <- compact_data_no_fmd %>% 
  filter(publication_year >= 2006, publication_year <= 2019) %>% 
  summarise(n = n())

print(paste('Number of studies between 2006 and 2019:', n_studies_btwn_2006_2019))
```

## Locations studied and connection of authors to the location

### Countries studied

```{r}
#Tally of how many times and percentage of times each COUNTRY has been studied
# countries_studied_by_count <- long_data_no_fmd %>% 
#     group_by(country_studied) %>%
#     distinct(title) %>% 
#     count(country_studied, name = 'num_of_studies', sort = T) %>% 
#     ungroup(country_studied) %>% 
#     mutate(country_studied = as_factor(country_studied), 
#            percentage = round(num_of_studies/sum(num_of_studies)*100, 1)
#            )

countries_studied_by_count <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>% 
    count(collab_type, country_studied, name = 'num_of_studies', sort = T) 

#'Modify the data frame to lump together all countries with less than some number of studies as 'other' and recalculate the percentage studied
countries_studied_by_count_aggregated <- countries_studied_by_count %>% 
  group_by(country_studied) %>%
  summarise(num_of_studies = sum(num_of_studies),
            .groups = 'drop'
            ) %>% 
  filter(country_studied %nin% c('none', 'west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  ungroup() 


#Print the results
knitr::kable(countries_studied_by_count_aggregated, 
             caption = 'Number of studies per country'
             )
``` 



### Continents studied
```{r echo=FALSE}
continents_studied_by_count <- countries_studied_by_count %>% 
  mutate(continent_studied = countrycode(country_studied, 
                                          destination = 'continent',
                                          origin = 'iso2c',
                                          nomatch = NULL #if there is no match, keep the original
                                          )
         ) %>% 
  mutate(continent_studied = str_to_lower(continent_studied)
         ) %>% 
  mutate(continent_studied = case_when(str_detect(continent_studied, 'asia') == T ~ 'asia', 
                                         str_detect(continent_studied, 'africa') == T ~ 'africa', 
                                         TRUE ~ continent_studied
                                       )
           ) %>% 
  group_by(continent_studied) %>%
  summarise(num_of_studies = sum(num_of_studies), 
            percentage = sum(percentage),
            .groups = 'drop'
            ) %>% 
  arrange(desc(percentage))
``` 




### Country most studied
```{r echo=FALSE}
countries_studied_most <- countries_studied_by_count %>%
    filter(country_studied != 'none') %>%
    filter(num_of_studies == max(num_of_studies))

#Print the result as a table
knitr::kable(countries_studied_most,
             col.names = c('Country',
                           'Number of studies',
                           'Percentage'
                           ),
             caption = 'Most studied country'
             )
```

### Continent most studied
```{r echo=FALSE}

continent_studied_most <- continents_studied_by_count %>%
    dplyr::filter(continent_studied != 'none') %>%
    filter(num_of_studies == max(num_of_studies))


#Print the result as a table
knitr::kable(continent_studied_most,
             col.names = c('Continent',
                           'Number of studies',
                           'Percentage'
                           ),
             caption = 'Most studied continent'
             )
```

## Number of studies per continent
```{r echo=FALSE}
continents_studied <- continents_studied_by_count %>%
  filter(continent_studied != 'none') %>%
  filter(continent_studied != 'global') %>%
  filter(continent_studied != 'northern_hemisphere') %>%
  group_by(continent_studied) %>%
  summarise(num_of_studies = sum(num_of_studies )) %>%
  ungroup() %>%
  mutate(percentage = round(num_of_studies/sum(num_of_studies)*100, 1))

#Print table of results
knitr::kable(continents_studied_by_count %>% arrange(desc(percentage)),
             col.names = c('Continent',
                           'Number of papers',
                           'Percentage'
                           ),
             caption = 'Number/percent of studies by continent'
             )

```



```{r echo=FALSE, include=FALSE}
# Load the cleaned data 
long_data <- read_rds('../data/final_data/cleaned_data/review_data_long_cleaned.rds')

#The cleaned data without citations
compact_data_init <- read_rds('../data/final_data/cleaned_data/review_data_compact_cleaned.rds')

compact_data <- compact_data_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

# The data with citations 
compact_data_with_citation_keys_init <- read_rds(normalizePath('../data/final_data/cleaned_data/compact_data_with_citation_keys_cleaned.rds')) 

compact_data_with_citation_keys <- compact_data_with_citation_keys_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

# Remove the FMD papers (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd') 

compact_data_no_fmd <- compact_data %>% filter(disease != 'fmd') 

compact_data_with_citation_keys_no_fmd <- compact_data_with_citation_keys %>% filter(disease != 'fmd') 
```

### Interventions types (FMD)

```{r echo=FALSE}
intervention_categories_fmd <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  group_by(intervention_type, vax_modelled_with_non_vax) %>% 
  count(is_vax_effective, sort = TRUE) %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
  adorn_totals(where = 'col')


knitr::kable(intervention_categories_fmd)
```

## Vaccine impact categories (FMD)
```{r echo=FALSE}
vax_impact_categories_fmd <- compact_data %>%
  filter(disease == 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, 
        name = 'num_of_studies', 
        sort = T
        ) 

#Print the results
knitr::kable(vax_impact_categories_fmd %>% adorn_totals, 
             col.names = c('Vaccination is most impactful', 
                           'Number of studies'
                           ),
             caption = 'Studies\' conclusions about impact of vaccination (FMD)'
             )
```


# Citation tables

## Code availability
```{r echo=FALSE}

code_availability_citations <- compact_data_with_citation_keys %>% 
  filter(simulation_code_available == 'yes')

knitr::kable(code_availability_citations)
```

## Citation table for human diseases
```{r echo=FALSE}
vax_impact_categories_citations_no_fmd <- compact_data_with_citation_keys %>% 
  filter(disease != 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  select(title, publication_year, disease, objectives, is_vax_effective, bibtexkey) %>% 
  arrange(is_vax_effective) 

#Print the results
knitr::kable(vax_impact_categories_citations_no_fmd,
             caption = 'Studies\' conclusions about impact of vaccination (citations for human diseases)'
             )
```

## Citation table for FMD
```{r echo=FALSE}
vax_impact_categories_citations_fmd <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  select(title, publication_year, disease, objectives, is_vax_effective, bibtexkey) %>% 
  arrange(is_vax_effective) 

#Print the results
knitr::kable(vax_impact_categories_citations_fmd,
             caption = 'Studies\' conclusions about impact of vaccination (citations for human diseases)'
             )

```

# Exploration of the peak periods (2009-2011, 2015-2016, 2019)

## Citations of studies during 2009-2011
```{r echo=FALSE}
#Citation table of studies during 2009-2011 (Peak period)
studies_in_2009_2011_df <- compact_data_with_citation_keys_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  filter(publication_year >= 2009, publication_year <= 2011) %>%
  arrange(disease) %>%
  select(title, publication_year, disease, objectives, bibtexkey) %>% 
  mutate(disease = str_replace(disease, ' ', ' and '))

knitr::kable(studies_in_2009_2011_df, caption = 'Studies published during 2009-2011 (Peak period)')
```
### Number of studies during 2009-2011 per disease and objective type (Peak period)
```{r echo=FALSE}
n_studies_2009_2011_df <- studies_in_2009_2011_df %>%
  count(disease, objectives) %>% 
  arrange(disease, n) %>% 
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(both = 0, future = 0, past = 0))

knitr::kable(n_studies_2009_2011_df, caption = 'Studies published during 2009-2011 (Peak period)')
```

## Citation table of studies during 2015-2016 per disease and objective type (Peak period)
```{r echo=FALSE}
studies_between_2015_2016_df <- compact_data_with_citation_keys_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  filter(publication_year >= 2015, publication_year <= 2016) %>%
  arrange(disease, publication_year) %>%
  select(title, publication_year, disease, objectives, bibtexkey) %>% 
  mutate(disease = str_replace(disease, ' ', ' and '))

knitr::kable(studies_between_2015_2016_df, caption = 'Studies published in 2015-2016 (Peak period)')
```

### Number of studies from 2015-2016 per disease and objective type (Peak period)
```{r echo=FALSE}
n_studies_between_2015_2016_df <- studies_between_2015_2016_df %>% 
  count(disease, objectives) %>% 
  arrange(desc(n))  %>% 
  pivot_wider(names_from = objectives, values_from = n) %>% 
  replace_na(list(both = 0, future = 0, past = 0)) 

knitr::kable(n_studies_between_2015_2016_df, caption = 'Papers published in 2015-2016 (Peak period)')
```

## Citation table of studies in 2019 (Peak period)
```{r echo=FALSE}
studies_in_2019_df <- compact_data_with_citation_keys_no_fmd %>%
  filter(publication_year == 2019) %>%
  arrange(disease) %>%
  select(title, publication_year, disease, objectives, bibtexkey) %>% 
  mutate(disease = str_replace(disease, ' ', ' and '))

knitr::kable(studies_in_2019_df, caption = 'Studies published in 2019 (Peak year)')
```
### Number of studies in 2019 per disease and objective type (Peak period)
```{r echo=FALSE}
n_studies_in_2019_df <- studies_in_2019_df %>% 
  count(disease, objectives) %>% 
  arrange(n)  

knitr::kable(n_studies_in_2019_df, caption = 'Number of studies published in 2019 per disease and objective type (Peak year)')
```

### Total number of studies in 2019 (Peak period)
```{r echo=FALSE}
n_studies_in_2019 <- n_studies_in_2019_df %>%  
  summarise(n = n())

print(paste('Number of studies in 2019 (Peak year):', 
            n_studies_in_2019
            )
      )
```

## Number of diseases in the peak periods
```{r echo=FALSE}
peak_period_df <- bind_rows(studies_in_2009_2011_df, 
                            studies_between_2015_2016_df, 
                            studies_in_2019_df
                            ) %>% 
  arrange(disease, publication_year)

knitr::kable(peak_period_df , 
             caption = 'Studies published in peak period'
             )
```

```{r echo=FALSE}
n_studies_by_disease_and_objectives_peak_periods <- peak_period_df %>% 
  count(disease, objectives) 

knitr::kable(n_studies_by_disease_and_objectives_peak_periods, 
             caption = 'Number of studies per disease and objective in peak periods'
             )
```

```{r echo=FALSE}
total_studies_by_disease_peak_period_df <- peak_period_df %>% 
  count(disease) %>% 
  arrange(n)

knitr::kable(total_studies_by_disease_peak_period_df, 
             caption = 'Number of studies per disease in peak periods'
             )
```

## Influenza during 2009 - 2011 peaks
```{r echo=FALSE}
n_Influenza_studies_by_objectives_2009_2011 <- peak_period_df %>% 
  filter(publication_year >= 2009, publication_year <= 2011) %>%
  filter(disease == 'influenza') %>% 
  count(objectives) 

knitr::kable(n_Influenza_studies_by_objectives_2009_2011, 
             caption = 'Number of Influenza studies per objective type (2009 & 2011)'
             )
```
## Ebola in 2015-2016
```{r echo=FALSE}
n_ebola_studies_by_objectives_2015_2016 <- studies_between_2015_2016_df %>% 
  filter(disease == 'ebola') %>% 
  count(objectives) 

knitr::kable(n_ebola_studies_by_objectives_2015_2016, 
             caption = 'Number of Ebola studies per objective type (2015-2016)'
             )
```
### Studies that used both retrospective and prospective assessments in peak periods
```{r echo=FALSE}
knitr::kable(peak_period_df %>% filter(objectives == 'both'), 
             caption = 'Studies with retrospective and prospective assessments'
             )

```

# Miscellaneous Analyses

```{r echo=FALSE}
disease_studied_count <- long_data_no_fmd %>% 
  group_by(disease) %>% 
  distinct(title) %>% 
  count(disease, name = 'num_of_studies', sort = T) %>%
  ungroup() 

knitr::kable(disease_studied_count, col.names = c('Disease', 'Number of studies'))
```

### Disease most studied
```{r echo=FALSE}
disease_studied_most <- disease_studied_count %>% 
  filter(num_of_studies == max(num_of_studies)) 

knitr::kable(disease_studied_most)
```


### Number of studies per collaboration type

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_no_fmd %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Purely academic', 
                                         'Mixed')
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        ) %>% 
  pivot_wider(names_from = collab_type, values_from = total_publications)

knitr::kable(collap_types_by_year %>% clean_names() %>% adorn_totals(... = c(mixed, purely_academic)))

```


### Academic collaborations versus other non-academic collaborations (excluding purely academic collaborations).

```{r echo=FALSE}
academic_collabs_versus_others <- compact_data_no_fmd %>% 
  filter(author_affiliation_type != 'academic_institutions') %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, 2005, publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(str_detect(author_affiliation_type, 'academic_institutions'), 
                                         'Academic', 
                                         'Other'
                                         )
                                 )
         ) %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )

#The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- '1970-2005'

#plot the result
academic_collabs_trend <- ggplot(data = academic_collabs_versus_others) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)  
             ) +
  scale_fill_manual(values = c('tan1', 'turquoise1')) +
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels
                     ) +
  labs(title = 'Academic and other collaborations over time',
       x = 'Year',
       y = 'Proportion of publications',
       fill = 'Collaboration type' 
         ) +
  theme_minimal(base_size = 12)

plot(academic_collabs_trend)


#Number of each collaboration type

```

# Connection of authors to the studied locations 
```{r}
author_location_count_by_disease <- compact_data_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  mutate(author_in_country_studied = as_factor(author_in_country_studied), 
         disease = as_factor(disease)
         ) %>% 
  group_by(collab_type, disease) %>%
  distinct(title, .keep_all = TRUE) %>% 
  count(author_in_country_studied, sort = TRUE) %>% 
  complete(author_in_country_studied, fill = list(n = 0)) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0, not_applicable = 0)) %>% 
  mutate(total = not_applicable + no + yes) %>% 
  ungroup(collab_type) %>% 
  arrange(desc(total))


knitr::kable(author_location_count_by_disease)
```
