---
title: "Primary analysis for Outbreak response models of vaccine-preventable diseases in humans (1970-2019)"
author: "James Azam"
date: "01/06/2021"
output:
  word_document:
    keep_md: yes
    toc: yes
#  html_notebook
---

```{r echo=FALSE, message=FALSE, include=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('ggthemes')
library('mdthemes')
library('scales')
library('janitor')
library('forcats')
library('stringr')
library('countrycode')
library('bib2df')
library('conflicted')

#resolve package function conflicts
conflict_prefer('filter', 'dplyr')
conflict_prefer('select', 'dplyr')
conflict_prefer('kable', 'knitr')
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'jpeg')

# Sets the working directory for subsequent chunks, but not this chunk
#knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```



```{r echo=FALSE, include=FALSE}
# Load the cleaned data 
long_data <- read_rds('../data/final_data/cleaned_data/review_data_long_cleaned.rds')

#The cleaned data without citations
compact_data_init <- read_rds('../data/final_data/cleaned_data/review_data_compact_cleaned.rds')

compact_data <- compact_data_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

# The data with citations 
compact_data_with_citation_keys_init <- read_rds(normalizePath('../data/final_data/cleaned_data/compact_data_with_citation_keys_cleaned.rds')) 

compact_data_with_citation_keys <- compact_data_with_citation_keys_init %>% 
    mutate(vax_modelled = ifelse(str_detect(intervention_modelled, 'vaccination'), 
                               'yes', 
                               'no')
         ) %>% 
  mutate(intervention_type = case_when(
    vax_modelled == 'no' & vax_modelled_with_non_vax == 'no' ~ 'no_vax', #Vaccination was not modelled
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'no' ~ 'vax_alongside', #when vax was modelled but not for comparison as a single intervention
    vax_modelled == 'yes' & vax_modelled_with_non_vax == 'yes' ~ 'vax_single', #when vax was modelled as a single intervention
  )) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) %>% 
  mutate(year_aggreg = as.numeric(ifelse(publication_year < 2006, #An approach to aggregate the years before 2006 into one group because of the low number of papers
                                         2005, 
                                         publication_year
                                         )
                                  ), 
         collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'purely_academic', 
                                         'mixed')
                                 )
         )

# Remove the FMD papers (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd') 

compact_data_no_fmd <- compact_data %>% filter(disease != 'fmd') 

compact_data_with_citation_keys_no_fmd <- compact_data_with_citation_keys %>% filter(disease != 'fmd') 
```


# Exploratory Data Analysis

```{r echo=FALSE, fig.dim=c(10, 6)}

publications_per_year <- compact_data_no_fmd %>% 
  count(publication_year) 

publications_per_year_plot <- ggplot(data = publications_per_year) + 
  geom_bar(aes(x = publication_year, 
               y = n
               ), 
           stat = 'identity', 
           color = 'black',
           fill = 'tomato3'
           ) +
  scale_y_continuous(breaks = seq(0, 30, 2),
                     labels = seq(0, 30, 2)
                     ) +
  labs(x = '**Year of publication**', y = '**Number of papers**') +
  md_theme_minimal(base_size = 12)
  

plot(publications_per_year_plot)

```

## Modelling objectives

```{r echo=FALSE}
#count the number of studies per objective type
modelling_objectives_count <- compact_data_with_citation_keys_no_fmd %>% 
  group_by(collab_type) %>% 
  count(objectives, name = 'num_of_studies') %>% 
  pivot_wider(names_from = objectives, values_from = num_of_studies) %>% 
  adorn_totals(where = c('col', 'row'))

#Print a table of the results
knitr::kable(modelling_objectives_count,
             caption = 'Study objectives by collaboration type'
             )
```

## Outbreak types
```{r echo=FALSE}
#Outbreak types per collab type
outbreak_types_by_collab_type <- compact_data_with_citation_keys_no_fmd %>% 
  group_by(collab_type) %>% 
  count(outbreak_type, name = 'num_of_studies') %>% 
  pivot_wider(names_from = outbreak_type, values_from = num_of_studies) %>% 
  adorn_totals(where = c('col', 'row'))

#Print a table of the results
knitr::kable(outbreak_types_by_collab_type,
             caption = 'Study objectives by collaboration type'
             )
```


## Modelling objectives and outbreak type by collaboration types
```{r echo=FALSE}
objectives_and_outbreak_types <- compact_data_with_citation_keys_no_fmd %>% 
  group_by(collab_type) %>% 
  count(objectives, outbreak_type, name = 'num_of_studies', sort = TRUE) %>% 
  pivot_wider(names_from = objectives, 
              values_from = num_of_studies
              ) %>% 
  replace_na(list(future = 0, past = 0, both = 0)) %>% 
  adorn_totals(where = c('col', 'row'))

knitr::kable(objectives_and_outbreak_types, 
             caption = 'Number of studies per objective and outbreak type'
             ) 
```

## Diseases studied per collab type

```{r echo=FALSE}
#count the number of unique diseases studied
disease_studied_by_collab_type <- compact_data_no_fmd %>% 
  separate_rows(disease,  sep = ' ') %>% 
  group_by(disease) %>% 
  distinct(title, .keep_all = TRUE) %>% 
  count(disease, collab_type, sort = TRUE) %>%
  pivot_wider(names_from = collab_type, values_from = n) %>% 
  replace_na(list(purely_academic = 0, mixed = 0)) %>% 
  mutate(total = purely_academic + mixed) %>% 
  arrange(desc(total))

knitr::kable(disease_studied_by_collab_type)
```

## Locations studied and connection of authors to the location

### Countries studied

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Tally of how many times and percentage of times each COUNTRY has been studied
countries_studied_by_count <- long_data_no_fmd %>% 
    group_by(country_studied) %>%
    distinct(title) %>% 
    count(country_studied, name = 'num_of_studies', sort = T) %>% 
    ungroup(country_studied) %>% 
    mutate(country_studied = as_factor(country_studied), 
           percentage = round(num_of_studies/sum(num_of_studies)*100, 1)
           )

#'Modify the data frame to lump together all countries with less than some number of studies as 'other' and recalculate the percentage studied
countries_studied_by_count_aggregated <- countries_studied_by_count %>% 
  mutate(country_studied_aggreg = if_else(num_of_studies < 9, #from eye balling the raw counts
                                          'other', 
                                          as.character(country_studied)
                                          )
         ) %>% 
  group_by(country_studied_aggreg) %>%
  summarise(num_of_studies = sum(num_of_studies),
            percentage = sum(percentage),
            .groups = 'drop'
            ) %>% 
  arrange(desc(percentage)) %>% 
  mutate(country_studied_aggreg = as_factor(country_studied_aggreg))

#Print the results
knitr::kable(countries_studied_by_count_aggregated, 
             col.names = c('Country', 
                           'Number of studies', 
                           'Percentage'
                           ),
             caption = 'Number/percentage of studies per country'
             )
``` 



### Continents studied
```{r echo=FALSE, warning=FALSE, message=FALSE}
continents_studied_by_count <- countries_studied_by_count %>% 
  mutate(continent_studied = countrycode(country_studied, 
                                          destination = 'continent',
                                          origin = 'iso2c',
                                          nomatch = NULL #if there is no match, keep the original
                                          )
         ) %>% 
  mutate(continent_studied = str_to_lower(continent_studied)
         ) %>% 
  mutate(continent_studied = case_when(str_detect(continent_studied, 'asia') == T ~ 'asia', 
                                         str_detect(continent_studied, 'africa') == T ~ 'africa', 
                                         TRUE ~ continent_studied
                                       )
           ) %>% 
  group_by(continent_studied) %>%
  summarise(num_of_studies = sum(num_of_studies), 
            percentage = sum(percentage),
            .groups = 'drop'
            ) %>% 
  arrange(desc(percentage))

#Print the results
knitr::kable(continents_studied_by_count, 
             col.names = c('Continent', 
                           'Number of studies', 
                           'Percentage'
                           ),
             caption = 'Number/percentage of studies per continent'
             )
``` 



# Connection of authors to the studied locations 
```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6}
author_location_count_by_disease <- compact_data_no_fmd %>%
  separate_rows(disease, sep = ' ') %>% 
  mutate(author_in_country_studied = as_factor(author_in_country_studied), 
         disease = as_factor(disease)
         ) %>% 
  group_by(collab_type, disease) %>%
  distinct(title, .keep_all = TRUE) %>% 
  count(author_in_country_studied, sort = TRUE) %>% 
  complete(author_in_country_studied, fill = list(n = 0)) %>% 
  pivot_wider(names_from = author_in_country_studied, values_from = n) %>% 
  replace_na(list(yes = 0, no = 0, not_applicable = 0)) %>% 
  mutate(total = not_applicable + no + yes) %>% 
  ungroup(collab_type) %>% 
  arrange(desc(total))

author_location_count_by_disease_long <- author_location_count_by_disease %>% 
  select(-c(not_applicable, total)) %>% 
  pivot_longer(cols = c(yes, no), 
               names_to = 'author_in_country_studied',
               values_to = 'n'
               )


author_location_count_by_disease_plot <- ggplot(data = author_location_count_by_disease_long, 
       aes(x = reorder(disease, n), 
           y = n
           )
       ) + geom_bar(aes(fill = author_in_country_studied),
                    stat = 'identity',
                    position = position_fill(reverse = TRUE)
                    ) +
  labs(x = '**Disease**', y = '**Proportion of total studies**') +
  md_theme_minimal(base_size = 12) +
  theme(legend.position = 'top') +
  facet_wrap(~ collab_type) +
  coord_flip()

print(author_location_count_by_disease_plot)
```

# Primary objectives 

## Trend and distribution of interactions between collaboration types

### Number of studies per combinations of collaboration types

Note that we do not include FMD in all the primary analyses. We only use it for comparisons, where necessary.

```{r echo=FALSE}
#' Count by collaboration types and make a bar plot
collab_types_count <- compact_data_no_fmd %>%
    count(author_affiliation_type, sort = TRUE) %>% 
    mutate(author_affiliation_type = as_factor(author_affiliation_type))

knitr::kable(collab_types_count %>% arrange(desc(n)))
```

### Trend in collaboration over the years

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
    
collap_types_count_barplot <- collab_types_count %>%
    ggplot() + geom_bar(aes(x = reorder(author_affiliation_type, n),
                            y = n
                            ),
                        stat = 'identity',
                        color = 'black'
                        ) +
    scale_y_continuous(breaks = seq(0, 140, 20),
                       labels = seq(0, 140, 20)
                       ) +
    labs(title = 'Number of studies per collaboration type', 
         x = '', 
         y = 'Number of studies') + 
  coord_flip() +
  theme_minimal() +
  theme(legend.position = 'none')

plot(collap_types_count_barplot)

knitr::kable(collab_types_count, 
             col.names = c('Collaboration type', 'Number of publications'),
             caption = 'Number of publications per collaboration type'
             )

```

### Proportions of the total publications per year by collaboration type

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_no_fmd %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )


#The plot
# Modify the x-axis labels to account for the lumping of the years from 2005 and below
x_axis_labels <- as.character(seq(2005, 2020, 2))

x_axis_labels[1] <- '1970-2005'


collab_type_by_year_plot <- ggplot(data = collap_types_by_year) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill(reverse = TRUE)  
             ) +
  scale_fill_manual(values = c('tan1', 'turquoise1')) +
  scale_x_continuous(breaks = seq(2005, 2020, 2),
                     labels = x_axis_labels,
                     expand = c(0, 0)
                     ) +
  coord_flip() +
  labs(title = 'Publications per collaboration type over time',
       x = 'Year',
       y = 'Proportion of total publications in the year',
       fill = 'Collaboration type' 
         ) +
  theme_minimal(base_size = 12)

plot(collab_type_by_year_plot)

```



```{r echo=FALSE}
#' Count the number of each author affiliation type per year
collap_types_by_year <- compact_data_no_fmd %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        ) 

collap_types_by_year_plot <- ggplot(data = collap_types_by_year, 
                                    aes(x = year_aggreg, 
                                        y = total_publications
                                        ), 
                                    color = 'black') + 
  geom_bar(aes(fill = collab_type), 
           stat = 'identity', 
           position = 'dodge'
           ) + 
  scale_fill_manual(values = c('tan2', 'turquoise3')) + 
  labs(x = '**Year**', y = '**Total publications**') +
  md_theme_minimal(base_size = 12)

plot(collap_types_by_year_plot)


```

Even though non-academic collaborations increased over time, the share of total publications per year over time decreased.



# Intervention types and impact of vaccination

## Classification of interventions by inclusion/exclusion of vaccination

### Human diseases
```{r echo=FALSE}
intervention_categories_human_diseases <- compact_data_no_fmd %>% 
  group_by(collab_type, intervention_type, vax_modelled_with_non_vax) %>% 
  count(is_vax_effective, sort = TRUE) %>% 
  pivot_wider(names_from = vax_modelled_with_non_vax, values_from = n) %>% 
  replace_na(list(no = 0, yes = 0))


knitr::kable(intervention_categories_human_diseases)
```
### FMD

```{r echo=FALSE}
intervention_categories_fmd <- compact_data %>% 
  filter(disease == 'fmd') %>% 
  group_by(intervention_type, vax_modelled_with_non_vax) %>% 
  count(is_vax_effective, sort = TRUE) %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  pivot_wider(names_from = is_vax_effective, values_from = n) %>% 
  adorn_totals(where = 'col')


knitr::kable(intervention_categories_fmd)
```

## Intervention types
```{r echo=FALSE}
study_categories_by_vax_single <- intervention_categories_human_diseases %>% 
  filter(intervention_type == 'vax_single')
  
knitr::kable(study_categories_by_vax_single)
```


# Is vaccination the most effective when compared to other single interventions?

## Vaccine impact categories (human diseases)
```{r echo=FALSE}
vax_impact_categories_no_fmd <- compact_data_no_fmd %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, 
        name = 'num_of_studies', 
        sort = T
        ) 

#Print the results
knitr::kable(vax_impact_categories_no_fmd %>% adorn_totals, 
             col.names = c('Vaccination is most impactful', 
                           'Number of studies'
                           ),
             caption = 'Studies\' conclusions about impact of vaccination (Human diseases)'
             )
```

## Vaccine impact categories (FMD)
```{r echo=FALSE}
vax_impact_categories_fmd <- compact_data %>%
  filter(disease == 'fmd') %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, 
        name = 'num_of_studies', 
        sort = T
        ) 

#Print the results
knitr::kable(vax_impact_categories_fmd %>% adorn_totals, 
             col.names = c('Vaccination is most impactful', 
                           'Number of studies'
                           ),
             caption = 'Studies\' conclusions about impact of vaccination (FMD)'
             )
```


```{r echo=FALSE, eval=FALSE}
vax_impact_plot <- ggplot(data = vax_impact_categories) + 
  geom_bar(
    aes(x = is_vax_effective, 
        y = n,
        fill = is_vax_effective
        ),
    stat = 'identity',
    width = 0.25
    ) +
  scale_y_continuous(breaks = seq(0, 
                                  max(vax_impact_categories$n),
                                  2
                                  ),
                     labels = seq(0, 
                                max(vax_impact_categories$n),
                                2
                                )
                     ) +
  scale_x_discrete(breaks = c('no', 
                              'yes', 
                              'the_outcomes_were_inconclusive'
                              ), 
                   labels = c('no' = 'No', 
                              'yes' = 'Yes', 
                              'the_outcomes_were_inconclusive' = 'Inconclusive'
                              )
                   ) +
  scale_fill_discrete(name = 'Vaccine impact',
                      labels = c('No', 
                                 'Inconclusive',
                                 'Yes'
                      )
                      ) +
  labs(x = 'Vaccination is most impactful single intervention',
       y = 'Number of studies'
       ) +
  theme_minimal(base_size = 12)

plot(vax_impact_plot)

```


# Interventions

# Outcomes measured

# Model structure, parametrization & validation, Sensitivity analysis

# Use of data data use and data availability

# Code availability



