---
title: "Systematic review of outbreak response models"
author: "James Azam"
date: "01/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('ggthemes')
library('scales')
library('janitor')
library('forcats')
library('stringr')
```


```{r echo=FALSE, message=FALSE}
#' Load the script that cleans the data and saves it; 
#' this should be run each time, just in case something has changed in the data cleaning process;
#' THe script clears the environment after cleaning the data
source(normalizePath('./scripts/data_cleaning/orv_models_review_data_cleaning.R'))
```

```{r echo=FALSE}
#' Load the cleaned data 

long_data <- read_rds(normalizePath('./data/final_data/cleaned_data/review_data_long_cleaned.rds'))

compact_data <- read_rds(normalizePath('./data/final_data/cleaned_data/review_data_compact_cleaned.rds'))

# Remove the FMD papers (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd') 

compact_data_no_fmd <- compact_data %>% filter(disease != 'fmd') 


#' 1. Author-affiliation and country studied/contextual knowledge (Policy-making related questions)
pm_data <- long_data_no_fmd %>% 
  select(
        paper_title,
        publication_year,
        objectives,
        author_affiliation_type,
        disease,
        country_studied,
        author_in_country_studied,
        outbreak_type,
        modelling_timing,
        data_used,
        data_available
        )  

#' 2. Vaccine-impact (Questions related to the Science done)
#' Select the variables related to vaccination modelling
vax_impact_data <- compact_data_no_fmd %>% 
    select(paper_title, 
           publication_year,
           disease,
           outbreak_type,
           intervention_modelled,
           vax_modelled_with_non_vax,
           is_vax_effective,
           outcome_measured
    ) 


```


# Exploratory Data Analysis
1. Publications per year
2. Distribution of collaborations across types (aggregated)
3. Countries/settings being studied (author affiliation vs country studied)


What science is being inferred from models?

1. Impact of vaccination (human diseases vs FMD); analyse with and without Influenza
2. What diseases are being studied (to understand vaccine impact)
3. Which interventions are being studied?

```{r echo=FALSE}

pubs_per_year <- compact_data_no_fmd %>% 
  mutate(publication_year = as.numeric(publication_year)) %>% 
  count(publication_year) 

pubs_per_year_bar_plot <- ggplot(data = pubs_per_year) + 
  geom_bar(aes(x = publication_year, 
               y = n
               ), 
           stat = 'identity', 
           color = 'black',
           fill = 'tomato3') +
  scale_y_continuous(breaks = seq(0, 26, 2),
                     labels = seq(0, 26, 2)
                     ) +
  labs(x = 'Year of publication', y = 'Number of papers') +
  theme_minimal()

plot(pubs_per_year_bar_plot)

```

```{r echo=FALSE}
modelling_objectives_by_outbreak_type_df <- long_data_no_fmd %>% 
  select(paper_title, objectives, outbreak_type) 

modelling_objectives_by_outbreak_type <- modelling_objectives_by_outbreak_type_df %>%
  group_by(paper_title) %>% 
  distinct(objectives, outbreak_type) %>% 
  ungroup() %>% 
  group_by(objectives) %>% 
  count(objectives, outbreak_type)

knitr::kable(modelling_objectives_by_outbreak_type)

modelling_objectives_n_timing_plot <- ggplot(data = modelling_objectives_by_outbreak_type) + 
  geom_bar(aes(y = n,
               x = objectives,
               fill = outbreak_type
               ),
           stat = 'identity', 
           position = position_dodge(),
           width = 0.6
           ) +
  labs(y = 'Number of papers') +
  scale_x_discrete(labels = c('both' = 'Prospective + Retrospective', 
                              'future' = 'Prospective', 
                              'past' = 'Retrospective')
                   ) +
  theme_minimal()

plot(modelling_objectives_n_timing_plot)
```



```{r table2, echo=FALSE, results = 'asis'} 
#Year with the highest publications
max_pub_year <- compact_data_no_fmd %>% 
    count(publication_year, sort = T) %>% 
    filter(n == max(n)) 

#Tally of how many times each disease was studied during the period
disease_studied_count <- long_data_no_fmd %>% 
  group_by(paper_title) %>% 
  distinct(disease) %>% 
  count(disease, sort = T) %>%
  ungroup() %>%
  count(disease, sort = T) 

knitr::kable(disease_studied_count)

# DIsease most studied
disease_studied_most <- disease_studied_count %>% 
  filter(n == max(n)) 

print(disease_studied_most)


#Tally of how many times each country has been studied
countries_studied_count <- long_data_no_fmd %>% 
    group_by(paper_title) %>%
    distinct(country_studied) %>% 
    count(country_studied, sort = T) %>% 
    ungroup(paper_title) %>% 
    count(country_studied, sort = T) 

knitr::kable(countries_studied_count)
  
countries_studied_most <- countries_studied_count %>% 
    dplyr::filter(country_studied != 'none') %>% 
    filter(n == max(n)) 

knitr::kable(tibble(max_pub_year = max_pub_year$publication_year, 
                    disease_with_max_pubs = disease_studied_most$disease,
                    country_most_studied = countries_studied_most$country_studied
                    )
             )

```

# Primary objectives 

1. Is modelling being used to inform outbreak response policy-making? 
+ How are the results distributed according to collaboration type?

Note that we do not include FMD in all the primary analyses. We only use it for comparisons, where necessary.

```{r echo=FALSE}
# 1B. Create the collaboration types and reword the string NA
collab_type_data <- compact_data_no_fmd %>%
  select(c(paper_title, publication_year, author_affiliation_type, author_in_country_studied)) %>% 
  mutate(publication_year = as.numeric(publication_year), 
         author_in_country_studied = as.character(author_in_country_studied)) %>% 
  arrange(publication_year) %>%
  mutate(
    author_affiliation_type = str_replace_all(author_affiliation_type, ' ', ' + '),
    author_in_country_studied = if_else(author_in_country_studied == 'NA', 'not_applicable',
                                        author_in_country_studied
                                        )
    ) 



#' Count by collaboration types and make a bar plot
collab_types_count <- collab_type_data %>%
    count(author_affiliation_type) %>% 
    mutate(author_affiliation_type = as_factor(author_affiliation_type))

knitr::kable(author_affiliation_count)
```

+ What has been the trend in collaboration over the years?

```{r echo=FALSE}
    
collap_types_count_barplot <- collab_types_count %>%
    ggplot() + geom_bar(aes(x = reorder(author_affiliation_type, n),
                            y = n,
                            fill = factor(author_affiliation_type)
                            ),
                        stat = 'identity',
                        color = 'black'
                        ) +
    scale_y_continuous(breaks = seq(0, 140, 20),
                       labels = seq(0, 140, 20)
                       ) +
    labs(title = 'Number of studies per collaboration type', 
         x = '', 
         y = 'Number of studies') + 
  coord_flip() +
  theme_minimal() +
  theme(legend.position = 'none')

plot(collap_types_count_barplot)

```

```{r echo=FALSE}
#' Count the number of each author affiliation type per year (including FMD)
collap_types_by_year <- compact_data_no_fmd %>% 
  arrange(publication_year) %>%
  mutate(collab_type = as_factor(if_else(author_affiliation_type == 'academic_institutions', 
                                         'Academic', 
                                         'Other collabs')
                                 )
  ) %>% 
  count(publication_year, 
        collab_type, 
        name = 'total_publications'
        )


collap_types_by_year_mod <- collap_types_by_year %>% 
  mutate(year_aggreg = as_factor(ifelse(publication_year < 2005, '<2005', publication_year)))

collap_types_by_year_mod_count <- collap_types_by_year_mod %>% 
  count(year_aggreg, 
        collab_type, 
        name = 'total_publications'
        )
knitr::kable(collap_types_by_year_mod_count)


```

A bar chart with each bar representing a proportion of the total publications for that year.

```{r echo=FALSE}

collab_type_by_year_plot <- ggplot(data = collap_types_by_year_mod_count) + 
  geom_bar(aes(x = year_aggreg,
               y = total_publications,
               fill = collab_type
               ),
           color = 'black',
           stat = 'identity',
           position = position_fill()  
             ) +
  scale_fill_manual(values = c('tan1', 'turquoise1')) +
  # scale_x_continuous(breaks = seq(1971, 2020, 5),
  #                    labels = seq(1971, 2020, 5)
  #                    ) +
  labs(title = 'Proportion of publications by collaboration type',
       x = 'Year',
       y = 'Proportion of publications',
       fill = 'Collaboration type' 
         ) +
  theme_minimal()

plot(collab_type_by_year_plot)

```

2. Is vaccination the most effective when compared to other single interventions?


```{r echo=FALSE}
#count the impact categories
vax_impact_categories <- compact_data_no_fmd %>% 
  filter(vax_modelled_with_non_vax == 'yes') %>% 
  count(is_vax_effective, sort = T) 

```


```{r echo=FALSE}
vax_impact_plot <- ggplot(data = vax_impact_categories) + 
  geom_bar(
    aes(x = is_vax_effective, 
        y = n,
        fill = is_vax_effective
        ),
    stat = 'identity',
    width = 0.25
    ) +
  scale_y_continuous(breaks = seq(0, 
                                  max(vax_impact_categories$n),
                                  2
                                  ),
                     labels = seq(0, 
                                max(vax_impact_categories$n),
                                2
                                )
                     ) +
  scale_x_discrete(breaks = c('no', 
                              'yes', 
                              'the_outcomes_were_inconclusive'
                              ), 
                   labels = c('no' = 'no', 
                              'yes' = 'yes', 
                              'the_outcomes_were_inconclusive' = 'inconclusive'
                              )
                   ) +
  scale_fill_discrete(name = 'Vaccine impact',
                      labels = c('no', 
                                 'inconclusive',
                                 'yes'
                      )
                      ) +
  labs(x = 'Vaccine is most impactful',
       y = 'Number of studies'
       ) +
  theme_minimal(base_size = 12)

plot(vax_impact_plot)

```

How many times each disease has been studied.

```{r}
#count the number of unique diseases studied
disease_studied <- long_data_no_fmd %>% 
    group_by(paper_title) %>% 
    distinct(disease) %>% 
    count(disease) %>% 
    ungroup() %>% 
    group_by(disease) %>% 
    count(disease) %>% 
    arrange(desc(n)) %>% 
    ungroup() %>% 
    mutate(disease = as_factor(disease))


diseases_plot <- ggplot(data = disease_studied) + 
    geom_bar(aes(x = disease, 
                 y = n
                 ), 
             stat = 'identity',
             fill = 'dodgerblue3',
             color = 'black'
             ) + 
    scale_y_continuous(breaks = seq(0, 140, 10),
                       labels = seq(0, 140, 10)
                       ) +
    coord_flip(expand = FALSE) +
    labs(x = 'Disease', y = 'Number of studies') +
    theme_minimal(base_size = 12)


plot(diseases_plot)
```


How many times each intervention type has been modelled.

```{r}
#count the number of unique interventions modelled
intervention_count <- long_data_no_fmd %>% 
    group_by(paper_title) %>% 
    distinct(intervention_modelled) %>% 
    count(intervention_modelled) %>% 
    ungroup() %>% 
    group_by(intervention_modelled) %>% 
    count(intervention_modelled) %>% 
    arrange(desc(n)) %>% 
    ungroup() %>% 
    mutate(intervention_modelled = as_factor(intervention_modelled))

intervention_count_plot <- ggplot(data = intervention_count) + 
  geom_bar(aes(x = intervention_modelled, 
               y = n), 
           stat = 'identity',
           fill = 'darkslategray2',
           color = 'black'
           ) +
  coord_flip() +
  theme_minimal()

plot(intervention_count_plot)

```

How many times each outcome/metric has been used

```{r}
#count the number of unique outcomes measured
outcome_count <- long_data_no_fmd %>% 
    group_by(paper_title) %>% 
    distinct(outcome_measured) %>% 
    count(outcome_measured) %>% 
    ungroup() %>% 
    group_by(outcome_measured) %>% 
    count(outcome_measured) %>% 
    arrange(desc(n)) %>% 
    ungroup() %>% 
    mutate(outcome_measured = as_factor(outcome_measured))

outcome_count_plot <- ggplot(data = outcome_count) + 
  geom_bar(aes(x = outcome_measured, 
               y = n), 
           stat = 'identity',
           fill = 'darkslategray4',
           color = 'black'
           ) +
  coord_flip() +
  theme_minimal()

plot(outcome_count_plot)


```

